Paper figures (Simulation)

### 2023-06-21 analysis/Paper_figures_simulation.Rmd
low PVE simulation bug fix
fix fusion perm

### Comparison with other methods
Bar plot: each bar shows the number of genes, colored by causal status. Use a different color for each method. The method and cut off values:
* ctwas: PIP 0.8
* FUSION fdr: 0.05
* FUSION bonferroni: 0.05
* FUSION permutation fdr: 0.05
* COLOC PP4: 0.8
* FOCUS PIP: 0.8
* SMR FDR: 0.05, updated using ensemble ID
* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05, updated using ensemble ID

@@

    } else if  (method == "fusionbon"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]
        }
    } else if  (method == "fusionperm"){
        res <- res[res$PERM.N!=0,]
        res$FDR <- p.adjust(res$PERM.PV, method = "fdr")
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]
        }
    } else if  (method == "coloc"){
        res <- res[!is.na(res$COLOC.PP4),]
        res <- res[order(res$COLOC.PP4, decreasing = T),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
         res <- res[res$COLOC.PP4 > cut,]
        }
    } else if (method == "focus"){
        res <- res[res$mol_name != "NULL",]
        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
        res <- res[order(res$pip, decreasing = T),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$pip > cut, ] 
        }
    }  else if (method == "SMR"){
        res <- as.data.frame(res)
        res$probeID <- sapply(res$Gene, function(x){unlist(strsplit(x, "[.]"))[1]})
        res <- res[res$probeID %in% weights$ENSEMBL_ID,]
        res <- res[sapply(res$p_HEIDI > 0.05, isTRUE),]
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
        res <- res[order(res$FDR, decreasing = F),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
           res <- res[res$FDR < cut,]
        }
    } else if (method == "MR-JTI"){
        res$ifcausal <- ifelse(res$variable %in% cau[[i]], 1, 0)
        res <- res[order(abs(res$beta), decreasing = T),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$CI_significance=="sig",] 
        }
    } else if (method == "PMR-Egger"){
        res <- as.data.frame(res)
        res$ifcausal <- ifelse(res$gene_id %in% cau[[i]], 1, 0) 
        res <- res[!is.na(res$causal_pvalue), ]
        res <- res[order(res$causal_pvalue, decreasing = F),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[sapply(res$causal_pvalue < cut/sum(!is.na(res$causal_pvalue)), isTRUE),]
        }
    } else{
      stop("no such method")
    }

@@

```{r ncausalplot, fig.width= 10, fig.height= 4}
colset = c("#ebebeb", "#ffffb3", # FOCUS
           "#ebebeb", "#8dd3c7", # Fusion
           "#ebebeb", "palegreen", # Fusion-permutation
           "#ebebeb", "#CC79A7", # MR-JTI
           "#ebebeb", "goldenrod", #PMR-Egger
           "#ebebeb", "#87CEFA", # SMR
           "#ebebeb", "#fb8072", # cTWAS
           "#ebebeb", "#bebada") # coloc
f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,225), main = "high gene PVE")
f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,225), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)
```

>>>>

### Comparison with other methods
Bar plot: each bar shows the number of genes, colored by causal status. Use a different color for each method. The method and cut off values:
* ctwas: PIP 0.8
* FUSION fdr: 0.05
* FUSION bonferroni: 0.05
* FUSION permutation: bonferroni < 0.05 then use empirical permutation p < 0.01
* COLOC PP4: 0.8
* FOCUS PIP: 0.8
* SMR FDR: 0.05, updated using ensemble ID
* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05, updated using ensemble ID

@@

    } else if  (method == "fusionbon"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]
        }
    } else if  (method == "fusionperm"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[res$PERM.N >0 & res$PERM.PV < 0.01,]
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]
        }
    } else if  (method == "coloc"){
        res <- res[!is.na(res$COLOC.PP4),]
        res <- res[order(res$COLOC.PP4, decreasing = T),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
         res <- res[res$COLOC.PP4 > cut,]
        }
    } else if (method == "focus"){
        res <- res[res$mol_name != "NULL",]
        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
        res <- res[order(res$pip, decreasing = T),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$pip > cut, ] 
        }
    }  else if (method == "SMR"){
        res <- as.data.frame(res)
        res$probeID <- sapply(res$Gene, function(x){unlist(strsplit(x, "[.]"))[1]})
        res <- res[res$probeID %in% weights$ENSEMBL_ID,]
        res <- res[sapply(res$p_HEIDI > 0.05, isTRUE),]
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
        res <- res[order(res$FDR, decreasing = F),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
           res <- res[res$FDR < cut,]
        }
    } else if (method == "MR-JTI"){
        res$ifcausal <- ifelse(res$variable %in% cau[[i]], 1, 0)
        res <- res[order(abs(res$beta), decreasing = T),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$CI_significance=="sig",] 
        }
    } else if (method == "PMR-Egger"){
        res <- as.data.frame(res)
        res$ifcausal <- ifelse(res$gene_id %in% cau[[i]], 1, 0) 
        res <- res[!is.na(res$causal_pvalue), ]
        res <- res[order(res$causal_pvalue, decreasing = F),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[sapply(res$causal_pvalue < cut/sum(!is.na(res$causal_pvalue)), isTRUE),]
        }
    } else{
      stop("no such method")
    }

@@

```{r ncausalplot, fig.width= 10, fig.height= 4}
colset = c("#ebebeb", "#ffffb3", # FOCUS
           "#ebebeb", "#8dd3c7", # Fusion
           "#ebebeb", "palegreen", # Fusion-permutation
           "#ebebeb", "#CC79A7", # MR-JTI
           "#ebebeb", "goldenrod", #PMR-Egger
           "#ebebeb", "#87CEFA", # SMR
           "#ebebeb", "#fb8072", # cTWAS
           "#ebebeb", "#bebada") # coloc
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,225), main = "high gene PVE")
f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,225), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)
```
